### 2020-12-16 docs/manuscript.tex
add Model backbone text

\usepackage{parskip}
\usepackage{microtype}
\usepackage{svg}
\usepackage{float}
\usepackage{placeins}
\usepackage[english]{babel} % Specify a different language here - english by default
\usepackage{listings}
\usepackage{fontspec}
\usepackage{hyperref}
% \usepackage{lipsum} % Required to insert dummy text. To be removed otherwise
\setmainfont{TeX Gyre Termes}
%----------------------------------------------------------------------------------------
%	COLUMNS
%----------------------------------------------------------------------------------------

@@

\subsection*{Model backbone}
\begin{itemize}
  \item logistic regression model, because easily extensible
  \item coefficients are interpretable as expected increase in log-odds of observing  a given replacement (also: average predictive comparisons to transform estimates to probability scale)
  \item probabilistic model allows us to perform posterior predictive checks
  \item disadvantages: none? maybe log-odds scale unintuitive. mention divide-by-4 rule
  \item  establish notation, start simple and expand on that? Maybe confusing because some things change as models get more complex 
\end{itemize}

>>>>

\usepackage{parskip}
\usepackage{microtype}
\usepackage{svg}
\usepackage{float}
\usepackage{placeins}
\usepackage[english]{babel} % Specify a different language here - english by default
\usepackage{listings}
\usepackage{fontspec}
\usepackage{hyperref}
\usepackage{amsmath}
% \usepackage{lipsum} % Required to insert dummy text. To be removed otherwise
\setmainfont{TeX Gyre Termes}
%----------------------------------------------------------------------------------------
%	COLUMNS
%----------------------------------------------------------------------------------------

@@

\subsection*{Model backbone}

We chose a logistic regression model as a backbone for two reasons: First, because it is easily extensible and secondly, because the coefficients are interpretable as expected increases on the log-odds scale, identical to the familiar interpretation of standard logistic regression models.
We also chose to fit the model in a Bayesian fashion using Stan, mostly because it allows us to make use of the hierarchical structure in the data and the possibility to use posterior predictive checks as a way of model checking.
Consider the following notation:

\begin{align}
  y_{ij} & \sim \text{bernoulli}(\theta_{ij}) \\
  \theta_{ij} & = \text{inv\_logit}(\beta_{0_{j}} + X_{i}\beta_{\text{HLA}_{j}})
\end{align},

where \(y_{ij}\) is the binary encoded observation at sequence \(i\) for replacement \(j\), \(\theta_{ij}\) is the estimated probability that we observe replacement \(j\) in sequence \(i\), \(X_{i}\) is the binary encoded (row) vector of HLA annotations for sequence \(i\), and \(\beta_{\text{HLA}_{j}}\) is the (column) vector of HLA regression coefficients for replacement \(j\) and  \(\text{inv\_logit}\) is the logistic link function. Note that, if we observe N different amino acid states at a certain alignment position, each of those are considered in turn to be the "replacement".

The regression coefficients \(\beta_{\text{HLA}_{j}}\) are the main focus of interest, as they quantify the association between the occurrence of replacement \(j\) and each of the observed HLA alleles. They are in units of log-odds, so a regression coefficient of somewhere around 2 means that when comparing sequences with and without that particular HLA allele, the log-odds of observing replacement \(j\) in presence of that HLA allele is increased by 2.

Reasoning on the log-odds can sometimes be unintuitive. A useful approximation to interpret logistic regression coefficients on the probability scale is the so-called divide-by-4 rule, which means that a regression coefficient of 2 corresponds to an expected increase on the probability scale of about 2/4 = 50\%.
